Introducing Distributed Dynamic Data-intensive (D3) Science: Understanding Applications and Infrastructure

### 2016-02-14 3dpas_intro.tex
ongoing.

Applications must either use one particular distribution, or transform data on
input to some stages. In particular, infrastructural applications require a
specific distribution for the data storage, and thus the stages that read and
process the data will need to make this choice.

As seen, how application data is distributed depends upon which stage of an
application is being considered. In spite of such variation, {\it data topology}
can be used to identify the flow of data through the application, with nodes
representing execution units and edges representing communication features.
Additionally, how the application reacts to and handles dynamism is an important
factor and may also indicate that some aspects of the data topology are unknown.

% The overall shape of such a data connectivity pattern gives a sense
% of {\it distributedness}; the thickness of edges gives a sense of
% how data intensive the application is; and the change in the shape
% and thickness (and rate) and number of edges gives a sense of the
% dynamism of the application. \katznote{it seems odd that we describe
%   a picture here but don't actually have any such picture in the
%   paper.}

One aspect of managing data topology is ``data planning," which refers
to how the data is coordinated from ingest through analysis to
output. In all cases, there is a coordination of the data with code
used to analyze or transform it. In some cases these form clear,
integrated stages where data flows from one to another. In others the
way that data enters and flows through the application is more loosely
defined and the execution of code is independent. Some examples
include the ATLAS experiments, which have a single point of primary
data generation which fans out to multiple analysis sites followed by
mass exchange of derived data, and sensor grids, which have multiple
sources of data generation collected to a single point of analysis
followed by constrained point of reuse.

We observe that pipelines provide a useful abstraction to represent
the application stages identified in
\figurename~\ref{fig:figures_application-stages}.  Many workflow tools
also enable simple data capture and processing pipelines to be
defined, where each pipeline stage may range from a complete
application to a subroutine in Matlab (for instance).  It is useful to
note that pipelines may also be dynamically constructed and adapted
using context and monitoring information, as undertaken in the
Industrial Incident Notification and Response application studied in
this paper. In this instance, the actual pipeline stage to be invoked
is triggered by the availability of particular types of data, a
feature also relevant in the Power Grids application.  However, it is
useful to note that different stages of the pipeline receive different
emphases in different applications. Some applications must collect and
store all incoming data, with only minor pre-processing to clean the
data ahead of the main processing. For others, a major constraint is
to reduce the incoming data stream as quickly as possible, perhaps
discarding a large fraction of the data before any processing
occurs. Reuse of pipeline stages across different applications must
therefore be undertaken with care.

>>>>

% Applications must either use one particular distribution, or transform data on
% input to some stages. In particular,

% Infrastructural applications require a specific distribution for the data
% storage, and thus  stages that read and process the data will need to make
% this choice.  As seen,

{\it Data topology} can be used to identify the flow of data through the
application, with nodes representing execution units and edges representing
communication features.  One aspect of managing data topology is ``data
planning," which refers to how the data is coordinated from ingest through
analysis to output. In all cases, there is a coordination of the data with code
used to analyze or transform it. In some cases these form clear, integrated
stages where data flows from one to another. In others the way that data enters
and flows through the application is more loosely defined and the execution of
code is independent. Some examples include the ATLAS experiments, which have a
single point of primary data generation which fans out to multiple analysis
sites followed by mass exchange of derived data, and sensor grids, which have
multiple sources of data generation collected to a single point of analysis
followed by constrained point of reuse.

% We observe that pipelines provide a useful abstraction to represent
% the application stages identified in
% \figurename~\ref{fig:figures_application-stages}.  Many workflow tools
% also enable simple data capture and processing pipelines to be
% defined, where each pipeline stage may range from a complete
% application to a subroutine in Matlab (for instance).  It is useful to
% note that pipelines may also be dynamically constructed and adapted
% using context and monitoring information, as undertaken in the
% Industrial Incident Notification and Response application studied in
% this paper. In this instance, the actual pipeline stage to be invoked
% is triggered by the availability of particular types of data, a
% feature also relevant in the Power Grids application.  However, it is
% useful to note that different stages of the pipeline receive different
% emphases in different applications.

Some applications must collect and store all incoming data, with only minor
pre-processing to clean the data ahead of the main processing. For others, a
major constraint is to reduce the incoming data stream as quickly as possible,
perhaps discarding a large fraction of the data before any processing
occurs.
